State of the Practice for Lattice Boltzmann Method Software

### 2022-01-24 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Updates to Section 2.1 on Qualities

We adopt software quality definitions from various researchers and subject
matter experts. Some of the definitions are from \cite{Smithetal2020}. The
following are the software quality definitions used in this exercise, along with
comments regarding their quantitative and qualitative measurement.  \wss{Update
when the Quality Definition of Qualities document is completed.}

\subsubsection{Installability}

Installability is measured by the effort required for the installation,
uninstallation or reinstallation of a software product in a specified
environment \citep{ISO/IEC25010, lenhard2013measuring}. A good measure of
installability correlates with scenarios when low or moderate effort is required
to gather and prepare software for its general use on a system for which it was
designed. In our case effort includes the time spent finding and understanding
the installation instructions, the person-time and resources spent performing
the installation procedure, and the absence or ease of overcoming system
compatibility issues. The ability to reasonably validate the installation
procedure and the ease of installation also have a positive effect on the
measure of installability.

\subsubsection{Correctness}

A software program is correct if it behaves according to its stated
specifications \citep{GhezziEtAl2003}. This requires that the specification is
available. Scientific software is unlikely to have a formal specification, since
this is not common practice in the field. As a consequence, the correctness of
software often cannot be measured directly. Despite an absent specification, the
correctness can indirectly be assessed by looking at the output and applying
domain knowledge. For our measurement of correctness we have assumed that the
following factors suggest correctness was considered: availability of a
requirements specification, reference to domain theory, and explicit use of
tools and/or techniques for building confidence of correctness, such as
documentation generators and software analysis tools.

\subsubsection{Verifiability}

Verifiability is measured by the extent to which a set of tests can be written
and executed to demonstrate that the delivered system meets the specification
\citep{sommerville}. Similarly to correctness, verifiability is correlated with
the availability of a specification and with reference to domain knowledge. A
good measure of verifiability is further correlated with the availability of
well written tutorials that include expected output, with software unit testing
documentation, and with evidence of continuous integration during the
development process. In our process we measure correctness and verifiability
together.

\subsubsection{Reliability}

Reliability is measured by the probability of failure-free operation of a
computer program in a specified environment for a specified time; that is,
reliability can be measured by the average time interval between two failures,
also known as the mean time to failure (MTTF) \citep{GhezziEtAl2003}
\citep{musa1987software}. Reliability is thus positively correlated with the
absence of errors during installation and use. Recoverability from errors also
improves reliability.

\subsubsection{Robustness}

Software possesses the characteristic of robustness if it behaves ``reasonably''
in two situations: i) when it encounters circumstances not anticipated in the
requirements specification; and ii) when the assumptions in its requirements
specification are violated \citep{boehm2007software, ghezzi1991fundamentals}. A
good measure of robustness correlates with a reasonable reaction to unexpected
input, including data of the wrong type, empty input, or missing files or links.
A reasonable reaction includes an appropriate error message and the ability to
recover the system.

\subsubsection{Performance}

Performance is measured by the degree to which a system or component
accomplishes its designated functions within given constraints, such as speed
(database response times, for instance), throughput (transactions per second),
capacity (concurrent usage loads), and timing (hard real-time demands)
\citep{IEEEStdGlossarySET1990, wiegers2003softreq}. In this state of the
practice assessment performance was not directly quantitatively measured.
Instead the documentation of each software package was observed for information
that alludes to a consideration of performance, such as parallelization tools.

\subsubsection{Usability}

Usability is measured by the extent to which a software product can be used by
specified users to achieve specified goals with effectiveness, efficiency, and
satisfaction in a specified context of use \citep{nielsonusability}. We assumed
that a high usability correlates with the presence of documentation,
including tutorials, manuals, and defined user characteristics, and user
support. Preferably the user support model has avenues to contact developers and
report issues.

\subsubsection{Maintainability}

A measure of maintainability is the effort with which a software system or
component can be modified to correct faults, improve performance or other
attributes, and satisfy new requirements \citep{IEEEStdGlossarySET1990,
boehm2007software}. In the current work maintainability is measured by the
quality of documentation artifacts, and the presence of version control and
issue tracking. These artifacts can greatly decrease the effort needed to modify
software. There are many documentation artifacts that can improve
maintainability, including user and developer manuals, specifications, README
files, change logs, release notes, publications, forums, and instructional
websites.

\subsubsection{Modifiability}

Modifiability refers to the ease with which stable changes can be made to a
system and the flexibility of the system to adopt such changes \citep{8016712}.
We did not directly measure modifiability. Instead, developers were asked in
interviews if they considered the ease of future changes when developing the
software packages, specifically changes to the structure of the system, modules
and code blocks. A follow up question is asked if any measures had been taken.

\subsubsection{Reusability}

Reusability refers to the extent to which components of a software package can
be used with or without adaptation in other software packages
\citep{kalagiakos2003non}. A good measure of reusability results from a large
number of easily reusable components. Increased software modularization, defined
as the presence of smaller components with well defined interfaces, is
important. For this state of the practice assessment, a good measure of
reusability correlates with an increased number of code files, and the
availability of API documentation.

\subsubsection{Understandability}

Understandability is measured by the capability of the software package to
enable the user to understand its suitability and function \citep{ISO9126}. It
is an artifact-dependent quality. Understandability is different for the
user-interface, source code, and documentation. In this state of the practice
analysis, understandability focuses on the source code. It is measured by the
consistency of a formatting style, the extend of modularization, the explicit
identification of coding standards, the presence of meaningful identifiers, and
clarity of comments.

\subsubsection{Traceability}

Traceability refers to the ability to link the software implementation and the
software artifacts, especially the requirement specification
\citep{McCallEtAl1977}. Similar to the quality of correctness, this requires the
availability of some form of specification. This quality refers to keeping track
of information as it changes forms or relates between artifacts. We did not
quantitatively measure traceability. Instead, developers were asked in
interviews how documentation fits into their development process.

\subsubsection{Visibility and Transparency}

Visibility and transparency refer to the extent to which all of the steps of a
software development process, and the current status of it, are conveyed clearly
\citep{ghezzi1991fundamentals}. In this state of the practice assessment a good
measure of visibility and transparency correlates with a well defined
development process, documentation of the development process and environment
and software package version release notes.

\subsubsection{Reproducibility}

Software achieves reproducibility if another developer can take the requirements
documentation and re-obtain the same software artifacts
\citep{BenureauAndRougier2017}. This includes the output of the software, where
the scientific results are compared between software implementations, or between
software implementations and manually calculated results. We measured
reproducibility qualitatively by asking developers if they have any concern that
their computational results won't be reproducible, and if they have taken any
steps to ensure reproducibility.

\subsubsection{Unambiguity}

Unambiguity refers to the extent to which two readers have similar
interpretations when reading software artifacts. In other words, artifacts are
unambiguous if, and only if, they only have one interpretation \citep{IEEE1998}.
This state of the practice assessment did not quantitatively measure
unambiguity, but we did ask developers if they think that the current
documentation can clearly convey all necessary knowledge to the users, and how
they achieved this or what improvements are needed to achieve it.

>>>>

We adopt software quality definitions from various researchers and subject
matter experts. The following are the software quality definitions used in this
exercise, along with comments regarding their measurement.

\begin{description}

	\item[Installability] Installability is measured by the effort required for
	the installation, uninstallation or reinstallation of a software product in
	a specified environment \citep{ISO/IEC25010, lenhard2013measuring}. A good
	measure of installability correlates with scenarios when low or moderate
	effort is required to gather and prepare software for its general use on a
	system for which it was designed. In our case effort includes the time spent
	finding and understanding the installation instructions, the person-time and
	resources spent performing the installation procedure, and the absence or
	ease of overcoming system compatibility issues. An installability score
	improves if the software has a means to validate the installation and
	instructions for uninstallation.

	\item[Correctness] A software program is correct if it behaves according to
	its stated specifications \citep{GhezziEtAl2003}. This requires that the
	specification is available. Scientific software is unlikely to have an
	explicit specification, since this is not common practice in the field. As a
	consequence, the correctness of software cannot often be measured directly.
	We assess correctness indirectly by looking for the following: availability
	of a requirements specification, reference to domain theory, and explicit
	use of tools and/or techniques for building confidence, such as
	documentation generators and software analysis tools.

	\item[Verifiability] An artifact (like code, or a theory manual) is
	verifiable if, and only if, every statement therein is verifiable. A
	statement is verifiable if, and only if, there exists some finite
	cost-effective process with which a person or machine can check that each
	statement is correct (adapted from recommended practice for requirements
	specifications \citep{IEEE1998}). Similarly to correctness, verifiability is
	correlated with the availability of a specification and with reference to
	domain knowledge. A good measure of verifiability is further correlated with
	the availability of well written tutorials that include expected output,
	with software unit testing, and with evidence of continuous integration. In
	our process we measure correctness and verifiability together.

	\item[Reliability] Reliability is measured by the probability of
	failure-free operation of a computer program in a specified environment for
	a specified time \citep{GhezziEtAl2003, musa1987software}. Reliability is
	thus positively correlated with the absence of errors during installation
	and use. Recoverability from errors also improves reliability.

	\item[Robustness] Software possesses the characteristic of robustness if it
	behaves ``reasonably'' in two situations: i) when it encounters
	circumstances not anticipated in the requirements specification; and ii)
	when the assumptions in its requirements specification are violated
	\citep{boehm2007software, ghezzi1991fundamentals}. A good measure of
	robustness correlates with a reasonable reaction to unexpected input,
	including data of the wrong type, empty input, or missing files or links.
	Examples of reasonable reaction include an appropriate error message and the
	ability to recover the system.

	\item[Performance] Performance is measured by the degree to which a system
	or component accomplishes its designated functions within given constraints,
	such as speed (database response times, for instance), throughput
	(transactions per second), capacity (concurrent usage loads), and timing
	(hard real-time demands) \citep{IEEEStdGlossarySET1990, wiegers2003softreq}.
	In this state of the practice assessment performance was not directly
	quantitatively measured. Instead the documentation of each software package
	was observed for information alluding to consideration of performance, such
	testing results, or parallelization instructions.

	\item[Usability] Usability is measured by the extent to which a software
	product can be used by specified users to achieve specified goals with
	effectiveness, efficiency, and satisfaction in a specified context of use
	\citep{nielsonusability}. We assumed that high usability correlates with
	the presence of documentation, including tutorials, manuals, defined
	user characteristics, and user support. Preferably the user support model
	will have avenues to contact developers and report issues.

	\item[Maintainability] A measure of maintainability is the effort with which
	a software system or component can be modified to correct faults, improve
	performance or other attributes, and satisfy new requirements
	\citep{IEEEStdGlossarySET1990, boehm2007software}. In the current work
	maintainability is measured by the quality of documentation artifacts, and
	the presence of version control and issue tracking. These artifacts can
	greatly decrease the effort needed to modify software. There are many
	documentation artifacts that can improve maintainability, including user and
	developer manuals, specifications, README files, change logs, release notes,
	publications, forums, and instructional websites.

	\item[Modifiability] Modifiability refers to the ease with which stable
	changes can be made to a system and the flexibility of the system to adopt
	such changes \citep{8016712}. We did not directly measure modifiability.
	Instead, developers were asked in interviews if they considered the ease of
	future changes when developing the software packages, specifically changes
	to the structure of the system, modules and code blocks. A follow up
	question is asked if any measures had been taken.

	\item[Reusability] Reusability refers to the extent to which components of a
	software package can be used with or without adaptation in other software
	packages \citep{kalagiakos2003non}. A good measure of reusability results
	from a large number of easily reusable components. Increased software
	modularization, defined as the presence of smaller components with well
	defined interfaces, is important. For this state of the practice assessment,
	we assumed that a good measure of reusability correlates with an increased
	number of code files, and the availability of API documentation.

	\item[Understandability] Understandability is measured by the capability of
	the software package to enable the user to understand its suitability and
	function \citep{ISO9126}. It is an artifact-dependent quality.
	Understandability is different for the user-interface, source code, and
	documentation. In this state of the practice analysis, understandability
	focuses on the source code. It is measured by the consistency of a
	formatting style, the extend of modularization, the explicit identification
	of coding standards, the presence of meaningful identifiers, and the clarity
	of comments.

	\item[Traceability] Traceability refers to the ability to link the software
	implementation and the software artifacts, especially the requirement
	specification \citep{McCallEtAl1977}. Similar to the quality of correctness,
	this requires the availability of some form of specification. This quality
	refers to keeping track of information as it changes forms or relates
	between artifacts. We did not quantitatively measure traceability. Instead,
	developers were asked in interviews how documentation fits into their
	development process.

	\item[Visibility and Transparency] Visibility and transparency refer to the
	extent to which all of the steps of a software development process, and the
	current status of it, are conveyed clearly \citep{ghezzi1991fundamentals}.
	In this state of the practice assessment a good measure of visibility and
	transparency correlates with a well defined development process,
	documentation of the development process and environment and software
	package version release notes.

	\item[Reproducibility] Software achieves reproducibility if another
	developer can take the requirements documentation and re-obtain the same
	software artifacts \citep{BenureauAndRougier2017}. This includes the output
	of the software, where the scientific results are compared between software
	implementations, or between software implementations and manually calculated
	results. We measured reproducibility qualitatively by asking developers if
	they have any concern that their computational results won't be
	reproducible, and if they have taken any steps to ensure reproducibility.

	\item[Unambiguity] Unambiguity refers to the extent to which two readers
	have similar interpretations when reading software artifacts. In other
	words, artifacts are unambiguous if, and only if, they only have one
	interpretation \citep{IEEE1998}. This state of the practice assessment did
	not quantitatively measure unambiguity, but we did ask developers if they
	think that the current documentation can clearly convey all necessary
	knowledge to the users, and how they achieved this or what improvements are
	needed to achieve it.

\end{description}
